---
title: 'ASSIGNMENT V'
author: "OMITI DENIS"
date: "2025-01-05"
output:
  word_document: default
  html_document:
    df_print: paged
---

```{r}
# Load necessary libraries
library(ggplot2)
library(lmtest)
library(car) # For Durbin-Watson test
library(MASS) # For additional diagnostics

# Load the data
eda_data <- read.csv("eda_data.csv")

# Fit the multiple linear regression model
model <- lm(Rating ~ avg_salary + age + Founded + Sector, data = eda_data)

# Display the model summary
summary(model)
```
```{R}
# Extract model coefficients
coefficients <- coef(model)
```

```{r}
# Print the fitted model equation
cat("Fitted Model Equation: Rating =", coefficients[1], "+", coefficients[2], "* avg_salary +", 
    coefficients[3], "* age +", coefficients[4], "* Founded +", coefficients[5], "* Sector\n")
```

#Interpretation of model results


R-squared is 0.4796 this indicates that 47.96% variation in ratings is explained by the explanatory variables in the fitted model that is, average salary,age,founded and sector, therefore the model provides moderate fit to the data. 

Adjusted R² penalizes the inclusion of unnecessary predictors. After accounting for the number of predictors, the model explains 46% of the variability in Rating. The small drop from R² indicates that most predictors are contributing meaningfully to the model.

Intercept: -0.554 indicates the baseline rating when all predictors are zero.
avg_salary: Coefficient (0.000133) is not significant (p-value = 0.825), implying no strong relationship with Rating.

age: Coefficient (0.000859) is marginally significant (p-value = 0.071).

Founded: Coefficient (0.000147) is significant (p-value = 0.006), suggesting older companies may have slightly higher ratings.

Sector: Most sector coefficients are highly significant, indicating that industry type strongly influences ratings. 

The intercept having the p_values 0.00541 less than 0.05 suggest that it is statistically significant in the model

The predictor variables founded and Sector also has p_values less than 0.05 indicating that they are statistically significant in the model.

F-statistic has Value: 24.37 (p-value < 2.2e-16) less than 0.05 indicates that the model is statistically significant, meaning at least one predictor is strongly associated with Rating.

Residual Standard Error (RSE) has Value: 0.5888, this indicates that the model's predictions deviate from the observed Rating by about 0.589 units on average.

```{r}
# Plot residuals vs fitted values
plot(model$fitted.values, model$residuals, 
     xlab = "Fitted values", ylab = "Residuals", 
     main = "Residuals vs Fitted")
abline(h = 0, col = "red")
```


#Interpretation


Residuals range from -4.919 to 4.142, with a median near zero, suggesting no major skewness.
```{r}

# Q-Q plot for normality
qqnorm(model$residuals)
qqline(model$residuals, col = "blue")
```


#Interpretation


Since the residuals lie close to the line in the Q-Q plot, it indicates that there is normality.
```{r}

# Perform Breusch-Pagan test for homoscedasticity
bptest(model)
```


#Interpretation


BP=36.529,p-value=0.1042: Residual variance is homoscedastic (constant variance), since the BP value is higher and the the p_value is greater than 0.05.
```{r}

# Generate diagnostic plots
par(mfrow = c(2, 2)) # Set plot layout
plot(model) # This will generate diagnostic plots
```


#interpretation


Residuals vs. Fitted: No clear pattern, indicating linearity assumption holds.

Q-Q Plot: Deviations suggest residuals deviate from normality.

Histogram of Residuals: Confirms lack of normality.

Scale-Location Plot: Shows some heteroscedasticity, but Breusch-Pagan suggests it's not severe.

Autocorrelation Plot: No major lags in residual autocorrelation.
```{r}
# Test assumptions of linear regression

# 1. Linearity assumption
plot(eda_data$avg_salary, eda_data$Rating, 
     main = "Scatter plot of avg_salary vs Rating",
     xlab = "avg_salary", ylab = "Rating")
abline(lm(Rating ~ avg_salary, data = eda_data), col = "red")
```
```{r}
# 2. Normality of residuals
hist(model$residuals, breaks = 20, main = "Histogram of Residuals", 
     xlab = "Residuals", col = "blue")
shapiro.test(model$residuals)  # Shapiro-Wilk test for normality
```


#Interpretation


W=0.882,p-value<2.2e−16: Residuals are not normally distributed.

```{r}
# 3. Homoscedasticity
plot(model$fitted.values, abs(model$residuals), 
     xlab = "Fitted values", ylab = "Absolute Residuals", 
     main = "Scale-Location Plot")
abline(h = 0, col = "red")
bptest(model)  # Breusch-Pagan test
```

```{r}
# 4. Independence of residuals
acf(model$residuals, main = "Autocorrelation of Residuals")
durbinWatsonTest(model)  # Durbin-Watson test for autocorrelation

```


#Interpretation


D-W statistic=1.985,p-value=0.852:  No significant autocorrelation in residuals since the D-W value 1.985 is very close to 2 with p-value greater than 0.05.
```{r}
# 5. Multicollinearity
vif_values <- vif(model)
print(vif_values)
```


#Interpretation


Since all the vif values are less than 5 it indicates that no multicollinearity that is there is no significant relationships among the predictors.

